#Approaches suggested by Laura, as defined in AshleyMHC.docx
```{r load_libraries, message=FALSE, warning=FALSE}
rm(list=ls())
options(warn=-1)
library(nlme)
library(car)
library(multcompView)
library(lsmeans)
library(rcompanion)
```
## Load HealthKit Step Count Data 
```{r load_and_filter_data, warning=FALSE}
healthkit_steps=data.frame(read.table(
  "health_kit_combined.steps.txt.regression",
  header=TRUE,sep='\t',stringsAsFactors = FALSE))

#Remove any step counts greater than 20k  
healthkit_steps=healthkit_steps[healthkit_steps$Value<20000,]

#restrict analysis to person-days that fall into one of study phases
phases=c("baseline","cluster","read_aha","stand","walk")
healthkit_steps=healthkit_steps[healthkit_steps$ABTest %in% phases,]
healthkit_steps$ABTest=factor(healthkit_steps$ABTest,levels=phases)

#phone only
healthkit_steps=healthkit_steps[healthkit_steps$WatchVsPhone=="phone",]

#subjects are random effects, generate factor from string
healthkit_steps$Subject=factor(healthkit_steps$Subject)

#keep track of weekIndex in addition to dayIndex
healthkit_steps$WeekIndex=ceiling(healthkit_steps$dayIndex/7) 
```
## 1) Assume there is no time effect 
```{r no_time_effect}
model = lme(Value ~ ABTest, 
            random = ~1|Subject,
            data=healthkit_steps,
            control=lmeControl(opt="optim"))
marginal = lsmeans(model, ~ ABTest )
print(summary(model))
print(marginal)
plot(cld(marginal,
    alpha   = 0.05, 
    Letters = letters,     ### Use lower-case letters for .group
    adjust  = "tukey"))     ###  Tukey-adjusted comparisons
#HISTOGRAM OF RESIDUALS
x = residuals(model)
plotNormalHistogram(x)
#plot(fitted(model),residuals(model))
```

## 2a) Test for a time effect within the baseline week using a day variable. 
```{r}
baseline=healthkit_steps[healthkit_steps$ABTest=="baseline",]
model = lme(Value ~ dayIndex, 
            random = ~1|Subject,
            data=baseline,
            control=lmeControl(opt="optim"))
marginal = lsmeans(model, ~ dayIndex )
print(summary(model))
print(marginal)
#HISTOGRAM OF RESIDUALS
x = residuals(model)
plotNormalHistogram(x)
```
## 2b) Test for a time effect within weeks 1-4 using a day variable 
```{r time_effect_weeks}
interventions=healthkit_steps[healthkit_steps$ABTest!="baseline",]
model = lme(Value ~ dayIndex, 
            random = ~1|Subject,
            data=interventions,
            control=lmeControl(opt="optim"))
marginal = lsmeans(model, ~ dayIndex )
print(summary(model))
print(marginal)

```
## 3) Assume the time effect is linear for all 5 weeks (use dayIndex or weekIndex as an explanatory variable in the model)
### using dayIndex 
```{r time_effect_linear_day}
model = lme(Value ~ ABTest+dayIndex, 
            random = ~1|Subject,
            data=healthkit_steps,
            control=lmeControl(opt="optim"))
marginal = lsmeans(model, ~ ABTest+ dayIndex )
print(summary(model))
print(marginal)
plot(cld(marginal,
    alpha   = 0.05, 
    Letters = letters,     ### Use lower-case letters for .group
    adjust  = "tukey"))     ###  Tukey-adjusted comparisons

```
### using weekIndex 
```{r time_effect_linear_week}
model = lme(Value ~ ABTest+WeekIndex, 
            random = ~1|Subject,
            data=healthkit_steps,
            control=lmeControl(opt="optim"))
marginal = lsmeans(model, ~ ABTest+ WeekIndex )
print(summary(model))
print(marginal)
plot(cld(marginal,
    alpha   = 0.05, 
    Letters = letters,     ### Use lower-case letters for .group
    adjust  = "tukey"))     ###  Tukey-adjusted comparisons
```
## 4a) Use the delta (intervention - baseline) as a dependent variable and look whether it differs from 0 on average. 
```{r delta_from_baseline}
library(dplyr,verbose=FALSE)
library(tidyr,verbose=FALSE)

#get a dataframem of mean-per-subject value for baseline as well as each intervention 
mean_intervention_vals=healthkit_steps %>%
  select(Subject,Date,ABTest,Value) %>%
  distinct(Subject,Date,ABTest,.keep_all = TRUE) %>%
  spread(ABTest,Value) %>%
  group_by(Subject) %>%
  summarise_all(funs(mean(., na.rm=TRUE)))

#get distributions of (intervention-baseline) for each intervention 
delta_cluster=mean_intervention_vals$cluster-mean_intervention_vals$baseline
delta_aha=mean_intervention_vals$read_aha-mean_intervention_vals$baseline 
delta_stand=mean_intervention_vals$stand - mean_intervention_vals$baseline 
delta_walk=mean_intervention_vals$walk - mean_intervention_vals$baseline 

#We want to see if mean of above 4 distributions differs from 0. 
#We first apply shapiro test to check if the values are normally distributed 
shapiro.test(delta_cluster)
shapiro.test(delta_aha)
shapiro.test(delta_stand)
shapiro.test(delta_walk)

#The shapiro test rejects the normal distribution in all 4 cases. 
#So we use Wilcoxon signed rank test to check for a shift from 0 
wilcox.test(delta_cluster, mu=0, conf.int = TRUE)
wilcox.test(delta_aha, mu=0, conf.int = TRUE)
wilcox.test(delta_stand, mu=0, conf.int = TRUE)
wilcox.test(delta_walk, mu=0, conf.int = TRUE)

#plot the distributions 
library(reshape2,verbose=FALSE)
library(ggplot2,verbose=FALSE)
deltas=data.frame(delta_aha,delta_walk,delta_cluster,delta_stand)
deltas=melt(deltas)
ggplot(data=deltas,aes(x=deltas$value,group=deltas$variable,color=deltas$variable,fill=deltas$variable))+
  geom_density(position="dodge",alpha=0.2)+
  theme_bw()
```
## 4b) 

## 4c) 

## 4d) 

## Repeat analysis for HealthKit Distance 

## Repeat analysis for MotionTracker 

## Repeat analysis for Watch Steps 

## Incorporating information from AWS analytics 
We have a subset of active subjects that did not receive stand prompts and daily step count reminders. What can we learn if we treat those as a separate group? 

